\documentclass[00-MASTER.tex]{subfiles}
\begin{document}
	
	%==================================================================================================== %
\section{Analyzing Influence in LME models}
 ``\textit{Influence}” is defined by \citet{schab} as ``the ability of a single or multiple data points, through their presence
	or absence in the data, to alter important aspects of the analysis, yield qualitatively different inferences, or
	violate assumptions of the statistical model". The goal of influence analysis is rather to identify influential cases and the manner in
	which they are important to the analysis. A consequence of this that cases may be to mark data
	points for deletion so that a better model fit can be achieved for the reduced data \citep{schab}.  
	
	% MOVE BACK TO START
	%  Outliers, for example, may be the most noteworthy data points in
	%  an analysis. They can point to a model breakdown and lead to development of a better model.
	
	
	\citet{schab} considers several important aspects of the use and implementation of influence measures in LME models. \textit{schabenberger} notes that it is not always possible to
	derive influence statistics necessary for comparing full- and reduced-data parameter estimates. 

	\citet{schab} describes a simple procedure for quantifying influence. Firstly a model should be fitted to the data, and
	estimates of the parameters should be obtained. The second step is that either single or multiple data points, specifically outliers,
	should be omitted from the analysis, with the original parameter estimates being updated. This is known as `\textit{leave one out \ leave k out}' analysis. The final step of the procedure is comparing the 	sets of estimates computed from the entire and reduced data sets to determine whether the absence of observations changed the
	analysis.
	
	%============================================================================================================================ %
	
	\section{Measures of Influence} 
	% DFBETA
	% DFFITS
	% PRESS
	The impact of an observation on a regression fitting can be determined by the difference between the estimated regression coefficient of a model with all observations and the estimated coefficient when the particular observation is deleted. DFBETA and DFFITS are well known measures of influenc. The measure DFBETA is the studentized value of this difference.	DFFITS is a statistical measured designed to a show how influential an observation is in a statistical model. DFFITS is closely related to the studentized residual.

		\begin{eqnarray}
		DFBETA_{a} &=& \hat{\beta} - \hat{\beta}_{(a)} \\
		&=& B(Y-Y_{\bar{a}}
		\\ DFFITS = {\widehat{y_i} -
		\widehat{y_{i(k)}} \over s_{(k)} \sqrt{h_{ii}}} 
	\end{eqnarray}
	
	The prediction residual sum of squares (PRESS) is an value associated with this calculation. When fitting linear models, PRESS can be used as a criterion for model selection, with smaller values indicating better model fits.
%\begin{displaymath}
%	PRESS = \sum(y-y^{(k)})^2
%
%\end{displaymath}
%	
%	\begin{itemize}
%		\item $e_{-Q} = y_{Q} - x_{Q}\hat{\beta}^{-Q}$
%		\item $PRESS_{(U)} = y_{i} - x\hat{\beta}_{(U)}$
%	\end{itemize}
%	
	
	\subsection{Influence Analysis for LME Models} %1.1.3
	The linear mixed effects model is a useful methodology for fitting a wide range of models. However, linear mixed effects models are known to be sensitive to outliers. \citet{CPJ} advises that identification of outliers is necessary before conclusions may be drawn from the fitted model.
	%-------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Influence in LME Models}
	
	Model diagnostic techniques, well established for classical models, have since been adapted for use with linear mixed effects models. Diagnostic techniques for LME models are inevitably more difficult to implement, due to the increased complexity.
	
	
	\subsection{Influence Statistics for LME models} %1.1.4
	Influence statistics can be coarsely grouped by the aspect of estimation that is their primary target:
	\begin{itemize}
		\item overall measures compare changes in objective functions: (restricted) likelihood distance (Cook and Weisberg 1982, Ch. 5.2)
		\item influence on parameter estimates: Cook's  (Cook 1977, 1979), MDFFITS (Belsley, Kuh, and Welsch 1980, p. 32)
		\item influence on precision of estimates: CovRatio and CovTrace
		\item influence on fitted and predicted values: PRESS residual, PRESS statistic (Allen 1974), DFFITS (Belsley, Kuh, and Welsch 1980, p. 15)
		\item outlier properties: internally and externally studentized residuals, leverage
	\end{itemize}
	%---------------------------------------------------------------------------%
	
	% % Beckman, Nachtsheim and Cook (1987) 
	\citet{Beckman} applied the \index{local influence}local influence method of Cook (1986) to the analysis of the LME model.
	While the concept of influence analysis is straightforward, implementation in mixed models is more complex. Update formulae for fixed effects models are available only when the covariance parameters are assumed to be known.
	
	If the global measure suggests that the points in $U$ are influential, the nature of that influence should be determined. In particular, the points in $U$ can affect the following
	
	\begin{itemize}
		\item the estimates of fixed effects,
		\item the estimates of the precision of the fixed effects,
		\item the estimates of the covariance parameters,
		\item the estimates of the precision of the covariance parameters,
		\item fitted and predicted values.
	\end{itemize}
	%---------------------------------------------------------------------------%
	\newpage
	\section{Influence analysis for LME Models} %1.7
	
	Likelihood based estimation methods, such as ML and REML, are sensitive to unusual observations. Influence diagnostics are formal techniques that assess the influence of observations on parameter estimates for $\beta$ and $\theta$. A common technique is to refit the model with an observation or group of observations omitted.
	
	\citet{west} examines a group of methods that examine various aspects of influence diagnostics for LME models.
	For overall influence, the most common approaches are the `likelihood distance' and the `restricted likelihood distance'.
	
	
\subsection{Influence Analysis for LME Models} %1.1.3
The linear mixed effects model is a useful methodology for fitting a wide range of models. However, linear mixed effects models are known to be sensitive to outliers. \citet{CPJ} advises that identification of outliers is necessary before conclusions may be drawn from the fitted model.

Standard statistical packages concentrate on calculating and testing parameter estimates without considering the diagnostics of the model.The assessment of the effects of perturbations in data, on the outcome of the analysis, is known as statistical influence analysis. Influence analysis examines the robustness of the model. Influence analysis methodologies have been used extensively in classical linear models, and provided the basis for methodologies for use with LME models.
Computationally inexpensive diagnostics tools have been developed to examine the issue of influence \citep{Zewotir}.
Studentized residuals, error contrast matrices and the inverse of the response variance covariance matrix are regular components of these tools.

Influence arises at two stages of the LME model. Firstly when $V$ is estimated by $\hat{V}$, and subsequent
estimations of the fixed and random regression coefficients $\beta$ and $u$, given $\hat{V}$.
	
	%-------------------------------------------------------------------------------------------------------------------------------------%
	
	
%	\newpage
%	\section{Covariance Parameters} %1.5
%	The unknown variance elements are referred to as the covariance parameters and collected in the vector $\theta$.
	% - where is this coming from?
	% - where is it used again?
	% - Has this got anything to do with CovTrace etc?
	%---------------------------------------------------------------------------%
	
	
	\subsection{INFLUENCE DIAGNOSTICS IN THE MIXED PROCEDURE}
	Key to the implementations of influence diagnostics in the MIXED procedure is the attempt to quantify
	influence, where possible, by drawing on the basic definitions of the various statistics in the classical linear
	model. 
	
	On occasion, quantification is not possible. Assume, for example, that a data point is removed
	and the new estimate of the G matrix is not positive definite. This may occur if a variance component
	estimate now falls on the boundary of the parameter space. Thus, it may not be possible to compute certain
	influence statistics comparing the full-data and reduced-data parameter estimates. However, knowing that
	a new singularity was encountered is important qualitative information about the data point’s influence on
	the analysis.
	
	The basic procedure for quantifying influence is simple:
	
	\begin{enumerate}
		\item Fit the model to the data and obtain estimates of all parameters.
		\item Remove one or more data points from the analysis and compute updated estimates of model parameters.
		\item Based on full- and reduced-data estimates, contrast quantities of interest to determine how the absence
		of the observations changes the analysis.
	\end{enumerate}
	We use the subscript (U) to denote quantities obtained without the observations in the set U. For example,
	%βb
	(U) denotes the fixed-effects “\textit{\textbf{leave-U-out}}” estimates. Note that the set U can contain multiple observations.
	
	
	%===================================================================================
	If the global measure suggests that the points in U are influential, you should next determine the nature of
	that influence. In particular, the points can affect
	\begin{itemize}
		\item the estimates of fixed effects
		\item the estimates of the precision of the fixed effects
		\item the estimates of the covariance parameters
		\item the estimates of the precision of the covariance parameters
		\item fitted and predicted values
	\end{itemize}
	
	It is important to further decompose the initial finding to determine whether data points are actually troublesome.
	Simply because they are influential “somehow”, should not trigger their removal from the analysis or
	a change in the model. For example, if points primarily affect the precision of the covariance parameters
	without exerting much influence on the fixed effects, then their presence in the data may not distort hypothesis
	tests or confidence intervals about $\beta$.
	%They will only do so if your inference depends on an estimate of the
	%precision of the covariance parameter estimates, as is the case for the Satterthwaite and Kenward-Roger
	%degrees of freedom methods and the standard error adjustment associated with the DDFM=KR option.
	
	%------------------------------------------------------------%
	
	
	\subsection*{Influence Analysis for LME Models} %1.1.3
	The linear mixed effects model is a useful methodology for fitting a wide range of models. However, linear mixed effects models are known to be sensitive to outliers. Christensen et al advises that identification of outliers is necessary before conclusions may be drawn from the fitted model.
	
	Standard statistical packages concentrate on calculating and testing parameter estimates without considering the diagnostics of the model. The assessment of the effects of perturbations in data, on the outcome of the analysis, is known as statistical influence analysis. Influence analysis examines the robustness of the model. 
	
	Influence analysis methodologies have been used extensively in classical linear models, and provided the basis for methodologies for use with LME models.
	
	Computationally inexpensive diagnostics tools have been developed to examine the issue of influence \citep{tewomir} (Zewotir).
	Studentized residuals, error contrast matrices and the inverse of the response variance covariance matrix are regular components of these tools.
	
	\subsection*{Influence Statistics for LME models} %1.1.4
	Influence statistics can be coarsely grouped by the aspect of estimation that is their primary target:
	\begin{itemize}
		\item overall measures compare changes in objective functions: (restricted) likelihood distance (Cook and Weisberg 1982, Ch. 5.2)
		\item influence on parameter estimates: Cook's  (Cook 1977, 1979), MDFFITS (Belsley, Kuh, and Welsch 1980, p. 32)
		\item influence on precision of estimates: CovRatio and CovTrace
		\item influence on fitted and predicted values: PRESS residual, PRESS statistic (Allen 1974), DFFITS (Belsley, Kuh, and Welsch 1980, p. 15)
		\item outlier properties: internally and externally studentized residuals, leverage
	\end{itemize}
%--------------------------------------------------------------%

	
	\subsection{Overall Influence}
	An overall influence statistic measures the change in the objective function being minimized. For example, in
	OLS regression, the residual sums of squares serves that purpose. In linear mixed models fit by
	\index{maximum likelihood} maximum likelihood (ML) or \index{restricted maximum likelihood} restricted maximum likelihood (REML), an overall influence measure is the \index{likelihood distance} likelihood distance [Cook and Weisberg ].
	
	
	\subsection{Influence Diagnostics: Basic Idea and Statistics} %1.1.2
	%http://support.sas.com/documentation/cdl/en/statug/63033/HTML/default/viewer.htm#statug_mixed_sect024.htm
	
	The general idea of quantifying the influence of one or more observations relies on computing parameter estimates based on all data points, removing the cases in question from the data, refitting the model, and computing statistics based on the change between full-data and reduced-data estimation. 
	
	
	\subsection{Cook's 1986 paper on Local Influence}%1.7.1
	Cook 1986 introduced methods for local influence assessment. These methods provide a powerful tool for examining perturbations in the assumption of a model, particularly the effects of local perturbations of parameters of observations.
	
	The local-influence approach to influence assessment is quitedifferent from the case deletion approach, comparisons are of
	interest.
	
	

\subsection{Quantifying Influence}  %1.1.6

The basic procedure for quantifying influence is simple as follows:

\begin{itemize}
	\item Fit the model to the data and obtain estimates of all parameters.
	\item Remove one or more data points from the analysis and compute updated estimates of model parameters.
	\item Based on full- and reduced-data estimates, contrast quantities of interest to determine how the absence of the observations changes the analysis.
\end{itemize}

\citet{cook86} introduces powerful tools for local-influence assessment and examining perturbations in the assumptions of a model. In particular the effect of local perturbations of parameters or observations are examined.


%---------------------------------------------------------------------------%
\newpage
\section{Extension of techniques to LME Models} %1.2

Model diagnostic techniques, well established for classical models, have since been adapted for use with linear mixed effects models.Diagnostic techniques for LME models are inevitably more difficult to implement, due to the increased complexity.

Beckman, Nachtsheim and Cook (1987) \citet{Beckman} applied the \index{local influence}local influence method of Cook (1986) to the analysis of the linear mixed model.

While the concept of influence analysis is straightforward, implementation in mixed models is more complex. Update formulae for fixed effects models are available only when the covariance parameters are assumed to be known.

If the global measure suggests that the points in $U$ are influential, the nature of that influence should be determined. In particular, the points in $U$ can affect the following

\begin{itemize}
	\item the estimates of fixed effects,
	\item the estimates of the precision of the fixed effects,
	\item the estimates of the covariance parameters,
	\item the estimates of the precision of the covariance parameters,
	\item fitted and predicted values.
\end{itemize}




	
\end{document}